### 2013-09-19 Publications/2013-ISBI14/01-abstract.tex
Added figure 1. Methods written down
Advancing with the ISBI 2014 paper

Several retrospective correction techniques are readily
available, generally inherited from \gls*{fmri} literature,
as those data are usually acquired with \gls*{epi} as well.
These correction methodologies have been evaluated and 
compared in physical phantoms designed for \gls*{fmri} 
experiments, but little is known about their impact
on \gls*{dmri} and more precisely on the resulting 
tractography.

In this work, we use a renowned digital phantom designed
for assessing tractography algorithms, and subject it to
a ``theoretically correct'' and plausible deformation. We
corrected data back, with several available techniques 
(namely fieldmap-based, reversed encoding-based, 
registration-based). Finally, we ranked the methods 
based on the geometrical accuracy, the dropout compensation, 
and the impact on the resulting tractography.

>>>>

Several retrospective correction techniques are readily
available. In this work, we use a renowned digital phantom 
designed for evaluating tractography. We subject the phantom 
to a ``theoretically correct'' and plausible deformation that
resembles the artifact under investigation. We correct data back, 
with three standard methodologies (namely fieldmap-based, 
reversed encoding-based, and registration-based). 
Finally, we rank the methods based on their geometrical 
accuracy, the dropout compensation, and their impact on 
the resulting tractography.

### 2013-09-19 Publications/2013-ISBI14/02-introduction.tex
Added figure 1. Methods written down
Advancing with the ISBI 2014 paper

In-vivo whole-brain connectivity analysis has been a
research topic of high interest for the last
5 years \cite{jones_studying_2008}. 
\Gls*{dmri} can be used to probe the
orientation of fiber bundles within the brain,
generally applying \gls*{epi} sequences.
After a signal reconstruction step, tractography
algorithms draw a map of sampled structure.
These maps can represent the actual trajectories
of fiber bundles (deterministic tractography,
\cite{basser_microstructural_2011}) or pixel-wise
probability of connection to a certain origin
(probabilistic tractography). Finally, the
information about these connections is collected
into a network matrix that can be subjected to
the so-called ``connectome analysis''.

However, such a complex workflow is prone to
flaws that have been detailedly reported 
\cite{jones_twenty-five_2010,jbabdi_tractography:_2011}.
In order to enable reliable and robust connectome
analyses of the brain, these challenging needs for
solutions must be addressed. One prominent source
of inaccuracy are the susceptibility-derived artifacts,
for which \Gls*{epi} schemes are specially sensitive.
As susceptibility changes at tissue interfaces,
so does the magnetic field. The consequence of the 
inhomogeneity of the field translates in a highly 
distorted imaged anatomy and a significant signal 
destruction of certain regions of the brain 
(e.g. the orbitofrontal lobe, for the proximity of the
air surrounding sinuses). This artifact has been
profoundly described, and a number of methodologies
have been proposed. First approaches to
retrospective correction \cite{jezzard_correction_1995},
relied on an extra \gls*{mri} acquisition (so-called
field mapping), that probes the inhomogeneity of the field.
A second theory-based breed of methodologies acquire a 
map of the \gls*{psf} of the \gls*{epi} readouts to correct 
for the artifact \cite{robson_measurement_1997}. 
Next generation of methodologies \cite{cordes_geometric_2000,
chiou_simple_2000}, make use of the acquisition of 
extra \gls*{epi} volumes 
with specific differences that enable correction, for instance
the opposed direction of the phase encoding gradients 
Finally, the last family of methodologies acquire an 
extra \gls*{t2} image, and use its anatomical correctness 
as reference to find the deformation map through nonlinear 
registration \cite{kybic_unwarping_2000,studholme_accurate_2000}.
Registration-based methods usually map
the \gls*{t2} image to the so-called baseline (B0) volume 
of \gls*{dmri}. The choice of \gls*{t2} is due to the strong
similarity of intensity distribution with the B0 of \gls*{dmri}
datasets. More recent works report extensions of the existing
techniques \cite{andersson_how_2003,holland_efficient_2010,
andersson_comprehensive_2012}, and combined approaches
\cite{zaitsev_point_2004}.

Even though aforementioned techniques for distortion correction
have been studied \cite{zeng_image_2002,wu_comparison_2008},
the lack of a gold-standard limits the possible benchmarking 
strategies. \cite{irfanoglu_effects_2012} opened the question
of the impact of the distortions on tractography. In this work,
we propose an evaluation framework using a detailed digital 
phantom designed for tractography. Using this framework,
we are able to compare several correction techniques and we
provide results related to their performance in the task
itself and their impact on subsequent tractography.

>>>>

In-vivo whole-brain connectivity analysis has been a
research topic of high interest for the last
5 years \cite{jones_studying_2008}. 
\Gls*{dmri} can be used to probe the
orientation of fiber bundles within the brain,
generally applying \gls*{epi} sequences.
After a signal reconstruction step 
(i.e. \cite{basser_microstructural_2011}), 
tractography algorithms draw a map of the sampled 
structures.
These maps can represent the actual trajectories
of fiber bundles (deterministic tractography) or
pixel-wise probability of connection to a certain origin
(probabilistic tractography). Finally, the
information about these connections is collected
into a network matrix that can be subjected to
the so-called ``connectome analysis''.

However, such a complex workflow is prone to
flaws that have been detailedly reported 
\cite{jones_twenty-five_2010,jbabdi_tractography:_2011}.
In order to enable reliable and robust connectome
analyses of the brain, some weak processing points must
be improved. One prominent source
of inaccuracy are the susceptibility-derived artifacts,
for which \Gls*{epi} schemes are specially sensitive.
As susceptibility changes at tissue interfaces,
so does the magnetic field. The consequence of the 
inhomogeneity of the field translates in a highly 
distorted imaged anatomy and a significant signal 
destruction of certain regions of the brain 
(e.g. the orbitofrontal lobe, for the proximity of the
air surrounding sinuses). This artifact has been
profoundly described, generally within \gls*{fmri}
literature because they are usually acquired with 
\gls*{epi} as \gls*{dmri}.

First approaches to
retrospective correction \cite{jezzard_correction_1995},
relied on an extra \gls*{mri} acquisition (so-called
field mapping), that probes the inhomogeneity of the field.
A second theory-based breed of methodologies acquire a 
map of the \gls*{psf} of the \gls*{epi} readouts to correct 
for the artifact \cite{robson_measurement_1997}. 
Next generation of methodologies \cite{cordes_geometric_2000,
chiou_simple_2000}, make use of the acquisition of 
extra \gls*{epi} volumes 
with specific differences that enable correction, for instance
the opposed direction of the phase encoding gradients 
Finally, the last family of methodologies acquire an 
extra \gls*{t2} image, and use its anatomical correctness 
as reference to find the deformation map through nonlinear 
registration \cite{kybic_unwarping_2000,studholme_accurate_2000}.
Registration-based methods usually map
the \gls*{t2} image to the so-called baseline (B0) volume 
of \gls*{dmri}. The choice of \gls*{t2} is due to the strong
similarity of intensity distribution with the B0 of \gls*{dmri}
datasets. More recent works report extensions of the existing
techniques \cite{andersson_how_2003,holland_efficient_2010,
andersson_comprehensive_2012}, and combined approaches
\cite{zaitsev_point_2004}.

Even though aforementioned techniques for distortion correction
have been studied \cite{zeng_image_2002,wu_comparison_2008},
the lack of a gold-standard limits the possible benchmarking 
strategies. \cite{irfanoglu_effects_2012} raised the question
of distortion-derived impacts in tractography.
In this work, we propose an evaluation framework
using a detailed digital phantom designed for tractography.
Using this framework,
we are able to compare several correction techniques and we
provide results related to their performance in the task
itself and their impact on subsequent tractography.

### 2013-09-19 Publications/2013-ISBI14/03-methods.tex
Added figure 1. Methods written down
Advancing with the ISBI 2014 paper

\begin{table}[h]
\caption{Simulated \gls*{mri} modalities}
\label{table:phantom}
\begin{center}
\begin{tabular}{c|c}
\hline
Modality & Descrition \\
\hline
\gls*{dti} & 30 directions, \\
\hline
\gls*{t1} & \\
\hline
\gls*{t2} & \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

   \begin{figure}[thpb]
      \centering
      \framebox{\parbox{3in}{Figure of the phantom}}
      %\includegraphics[scale=1.0]{figurefile}
      \caption{Original and distorted digital phantom}
      \label{fig:label}
   \end{figure}

\subsection{Evaluation workflows}

>>>>

\begin{table}[h]
\caption{Simulated \gls*{mri} modalities}
\label{table:phantom}
\begin{center}
\begin{tabular}{c|c}
\hline
Modality & Description \\
\hline
\gls*{dti} & 30 directions, \\
\hline
\gls*{t1} & \\
\hline
\gls*{t2} & \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

Fieldmap-based correction methodologies
\cite{jezzard_correction_1995} use a map
of the field in the scanner. More precisely, the
phase difference between two subsequent samplings
of the fieldmap. With that information, it is possible
to compute the theoretical displacements that each
voxel is to suffer, the so-called \gls*{vsm}. The
most prominent feature of this \gls*{vsm} is that all
the shifts have the same orientation (parallel to the
phase-encoding direction of the \gls*{epi}) and their
magnitude and direction depend on the \gls*{epi} 
gradient increments (or \emph{blips}), and the actual
phase difference at the voxel.

In order to create a realistic distortion, we started
by generating a synthetic phase-difference map 
consistent with the phantom.  We defined two regions 
of smooth phase inhomogeneity and computed the 
corresponding \gls*{vsm} using the tools distributed 
with the \texttt{FSL} package \cite{jenkinson_fsl_2012}.

Then, it is straightforward to use the \gls*{vsm} to 
produce the deformation of the original phantom data,
using the \texttt{fugue} tool included in
\cite{jenkinson_fsl_2012}. With the same procedure,
just setting the opposite encoding direction, we obtained
the corresponding reversed encoding ``acquisition'' of
the same phantom, in order to use it in evaluating the
second family of correction methods.

In this simple manner, we generated a full gold-standard
protocol, containing realistic \gls*{t1} and \gls*{t2}
at high resolution, \gls*{dti} data acquired in two
different phase encoding directions, and the ground-truth
\gls*{dti} data that is not available for real datasets.

\begin{figure}[thpb]
   \centering
   \includegraphics[width=0.8\columnwidth]{Fig01-Phantom}
   \caption{Original and distorted digital phantom (sagittal
   planes). A) T2-weighted image; B) undistorted B0 volume;
   C and D) distorted B0 volumes with opposed phase encoding 
   directions.}
   \label{fig:label}
\end{figure}

\subsection{Correction methods}

\paragraph{\Gls*{fmb}} As it was previously described,
this family of correction technologies infer a \gls*{vsm}
from the phase difference map. In this work, we use the
tool \texttt{fugue} \cite{jenkinson_fsl_2012} aforementioned.

\paragraph{\Gls*{reb}} We use the tool
\texttt{topup} from \cite{jenkinson_fsl_2012} to correct
using the B0 of a reverse-encoding simulation of the \gls*{dti}.
In this second case, the \gls*{vsm} is inferred from the
differences between the two corresponding B0. This technique
also uses least-squares-fitting to interpolate the undistorted
signal, and therefore provides with a robust dropout correction.

\paragraph{\Gls*{t2b}} For this last methodology
we fine-tuned the \texttt{ANTs} tool \cite{avants_ants:_2013},
in order that registering the B0 to the T2. Additionally,
undistorted images are corrected for dropout using the determinant
of the Jacobian of the deformation field.

\subsection{Evaluation}

\paragraph*{Proposed workflows}
We used \texttt{nipype} \cite{gorgolewski_nipype:_2011}
to conveniently integrate the correction methods in
comprehensive evaluation workflows:

\begin{itemize}
\item \textbf{Phantom deformation}: We release the pipeline
that generates the gold-standard, along with
distorted versions of the \glspl*{tpm} for
evaluation purposes.
\item \textbf{Correction workflow}: A high-level pipeline
integrates the three proposed correction methods,
ensuring that all the requirements for inputs are
met. As outputs this workflow produces the different
corrected versions of the phantom (\gls*{fmb}, \gls*{reb}
and \gls*{t2b}) alongside corrected versions of the
distorted \glspl*{tpm}.
\item \textbf{Reconstruction workflow}: This pipeline uses
the Diffusion Toolkit \cite{wang_diffusion_2007}, 
to produce deterministic tractography
from \gls*{dmri} data at the input. Therefore,
tractographies of original, and recovered data
are produced for comparison.
\item \textbf{Benchmarking workflow}: The final processing
element collects all the outputs from the previous
workflows and extracts the evaluation scores.
\end{itemize}
General-purpose 
workflows have been contributed to the upstream code
of the tool, under the branch \texttt{nipype.workflows.dmri}.
Testing-specific workflows are also publicly released.

\paragraph*{Evaluation scores}
In order to assess the geometrical correctness of the
recovered phantoms, we report overlap indices, specifically
the fuzzy Jaccard index as proposed in
\cite{crum_generalized_2006} weighted by the volumes of
labels. The recovered signal \emph{sanity} (dropout 
correction) is evaluated with correlation maps, averaged
along the time axis to take into account all the volumes
existing on the \gls*{dmri} data. This checkpoint
assumes that geometrical distortion is precisely
corrected. Finally, we report the impact on tractography
using the \gls*{frr}, that informs the number of recovered
fibers versus the gold-standard number for a specific
reconstruction method, and the error on the resulting
connectivity matrices.

### 2013-09-19 Publications/2013-ISBI14/04-results.tex
Added figure 1. Methods written down
Advancing with the ISBI 2014 paper

\section{MATH}

Before you begin to format your paper, first write and save the content as a separate text file. Keep your text and graphic files separate until after the text has been formatted and styled. Do not use hard tabs, and limit use of hard returns to only one return at the end of a paragraph. Do not add any kind of pagination anywhere in the paper. Do not number text heads-the template will do that for you.

Finally, complete content and organizational editing before formatting. Please take note of the following items when proofreading spelling and grammar:

\subsection{Abbreviations and Acronyms} Define abbreviations and acronyms the first time they are used in the text, even after they have been defined in the abstract. Abbreviations such as IEEE, SI, MKS, CGS, sc, dc, and rms do not have to be defined. Do not use abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}

\begin{itemize}

\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as Ò3.5-inch disk driveÓ.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ÒWb/m2Ó or Òwebers per square meterÓ, not Òwebers/m2Ó.  Spell out units when they appear in text: Ò. . . a few henriesÓ, not Ò. . . a few HÓ.
\item Use a zero before decimal points: Ò0.25Ó, not Ò.25Ó. Use Òcm3Ó, not ÒccÓ. (bullet list)

\end{itemize}

\subsection{Equations}

The equations are an exception to the prescribed specifications of this template. You will need to determine whether or not your equation should be typed using either the Times New Roman or the Symbol font (please no other font). To create multileveled equations, it may be necessary to treat the equation as a graphic and insert it into the text after your paper is styled. Number equations consecutively. Equation numbers, within parentheses, are to position flush right, as in (1), using a right tab stop. To make your equations more compact, you may use the solidus ( / ), the exp function, or appropriate exponents. Italicize Roman symbols for quantities and variables, but not Greek symbols. Use a long dash rather than a hyphen for a minus sign. Punctuate equations with commas or periods when they are part of a sentence, as in

$$
\alpha + \beta = \chi \eqno{(1)}
$$

Note that the equation is centered using a center tab stop. Be sure that the symbols in your equation have been defined before or immediately following the equation. Use Ò(1)Ó, not ÒEq. (1)Ó or Òequation (1)Ó, except at the beginning of a sentence: ÒEquation (1) is . . .Ó

\subsection{Some Common Mistakes}
\begin{itemize}

\item The word ÒdataÓ is plural, not singular.
\item The subscript for the permeability of vacuum ?0, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ÒoÓ.
\item In American English, commas, semi-/colons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ÒinsetÓ, not an ÒinsertÓ. The word alternatively is preferred to the word ÒalternatelyÓ (unless you really mean something that alternates).
\item Do not use the word ÒessentiallyÓ to mean ÒapproximatelyÓ or ÒeffectivelyÓ.
\item In your paper title, if the words Òthat usesÓ can accurately replace the word ÒusingÓ, capitalize the ÒuÓ; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ÒaffectÓ and ÒeffectÓ, ÒcomplementÓ and ÒcomplimentÓ, ÒdiscreetÓ and ÒdiscreteÓ, ÒprincipalÓ and ÒprincipleÓ.
\item Do not confuse ÒimplyÓ and ÒinferÓ.
\item The prefix ÒnonÓ is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ÒetÓ in the Latin abbreviation Òet al.Ó.
\item The abbreviation Òi.e.Ó means Òthat isÓ, and the abbreviation Òe.g.Ó means Òfor exampleÓ.

\end{itemize}

>>>>

\section{RESULTS AND DISCUSSION}

WIP: work in progess..

### 2013-09-19 Publications/2013-ISBI14/05-conclusions.tex
Added figure 1. Methods written down
Advancing with the ISBI 2014 paper

\section{USING THE TEMPLATE}

Use this sample document as your LaTeX source file to create your document. Save this file as {\bf root.tex}. You have to make sure to use the cls file that came with this distribution. If you use a different style file, you cannot expect to get required margins. Note also that when you are creating your out PDF file, the source file is only part of the equation. {\it Your \TeX\ $\rightarrow$ PDF filter determines the output file size. Even if you make all the specifications to output a letter file in the source - if your filter is set to produce A4, you will only get A4 output. }

It is impossible to account for all possible situation, one would encounter using \TeX. If you are using multiple \TeX\ files you must make sure that the ``MAIN`` source file is called root.tex - this is particularly important if your conference is using PaperPlaza's built in \TeX\ to PDF conversion tool.

\subsection{Headings, etc}

Text heads organize the topics on a relational, hierarchical basis. For example, the paper title is the primary text head because all subsequent material relates and elaborates on this one topic. If there are two or more sub-topics, the next level head (uppercase Roman numerals) should be used and, conversely, if there are not at least two sub-topics, then no subheads should be introduced. Styles named ÒHeading 1Ó, ÒHeading 2Ó, ÒHeading 3Ó, and ÒHeading 4Ó are prescribed.

\subsection{Figures and Tables}

Positioning Figures and Tables: Place figures and tables at the top and bottom of columns. Avoid placing them in the middle of columns. Large figures and tables may span across both columns. Figure captions should be below the figures; table heads should appear above the tables. Insert figures and tables after they are cited in the text. Use the abbreviation ÒFig. 1Ó, even at the beginning of a sentence.

\begin{table}[h]
\caption{An Example of a Table}
\label{table_example}
\begin{center}
\begin{tabular}{|c||c|}
\hline
One & Two\\
\hline
Three & Four\\
\hline
\end{tabular}
\end{center}
\end{table}

   \begin{figure}[thpb]
      \centering
      \framebox{\parbox{3in}{We suggest that you use a text box to insert a graphic (which is ideally a 300 dpi TIFF or EPS file, with all fonts embedded) because, in an document, this method is somewhat more stable than directly inserting a picture.
}}
      %\includegraphics[scale=1.0]{figurefile}
      \caption{Inductance of oscillation winding on amorphous
       magnetic core versus DC bias magnetic field}
      \label{figurelabel}
   \end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words rather than symbols or abbreviations when writing Figure axis labels to avoid confusing the reader. As an example, write the quantity ÒMagnetizationÓ, or ÒMagnetization, MÓ, not just ÒMÓ. If including units in the label, present them within parentheses. Do not label axes only with units. In the example, write ÒMagnetization (A/m)Ó or ÒMagnetization {A[m(1)]}Ó, not just ÒA/mÓ. Do not label axes with a ratio of quantities and units. For example, write ÒTemperature (K)Ó, not ÒTemperature/K.Ó

>>>>

\section{CONCLUSION}

This paper proposes an evaluation framework to
comprehensively analyze the impact of susceptibility
induced distortion on tractography from \gls*{dmri}
data. Inaccuracy on tractography leads to significant
errors in posterior connectivity analyses of the
whole-brain, specially in distortion affected regions.
We publicly release the framework and also contribute
with the evaluation of three widely-used available
correction methodologies. We conclude that the 
\gls*{reb} method resulted the first method in the 
proposed rankings (geometric correctness, signal loss
recovery and tractography impact), with the drawback
of demanding some extra \gls*{dmri} data.
